---
title: 'Ontario Democracy Under Pandemic: Explaining How the Ford Government Prevailed
  During COVID-19'
author: "Bianca Pokhrel, Cameron Fryer, Huining Qu, Zhuoqian Li"
date: "2020/10/09"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# Excecutive Summary
With COVID-19 rampages through Ontario, the PPC (Progressive Conservative Party) of Ontario, under the leadership of Premier Doug Ford, achieved initial success in containing the spread of the virus, and along with it, Ford’s support rate has increased. The survey used in the report is designed to analyze the impact of the pandemic and its relations with the approval rate of Premier Doug Ford and his party.

Our survey, in total of 12 questions, gives us an overview of COVID-19 impact, government performance rating, and voting intentions. The survey starts with socio-demographic questions(like gender and age) to ensure all segments of the Ontario population are properly represented. We have ensured to arrange the questions in a specific order so that the previous question has minimal impact on respondents' answers. Then it is administered using simple random sampling without replacement(SRSWOR) through dialing random Ontario residents. The survey result was then simulated by datasets from recently conducted polls online. The findings have demonstrated the percentage of residents impacted by the pandemic from each perspective. The general satisfaction with the Ford government’s response to the pandemic is around 77%.  Though the majority of the respondents indicated that there have been no changes in their work, financial situation, and mental health, it is extremely concerning that about 34% of the respondents have experienced a negative impact on their mental health. The result above indicates that more resources are mandatory to help Ontario residents experiencing mental health difficulties. Several significant findings come from respondents’ voting intentions. 47% of respondents would vote for the Conservative Party if a provincial election is being held today, which is a 5.3% increase from the 2018 election. Moreover,  21% of decided voters have changed their political stands based on the government's response to COVID.  Such findings suggest that a key to gain popularity from Ontarians is through taking proper action of the pandemic. As the Progressive Conservative Party of Ontario passes the halfway mark of its political term, the rising approval rate elevates the party’s electoral fortune in the 2020 election. 

We have taken a quantitative approach by surveying a large sample of randomly selected respondents. The main shortcoming of quantitative methods is that the interpretation is not discussed thus left to the respondent’s discretion. Our survey only focuses on preconceived issues(eg. Public health response, work, finance, mental health, vote intentions), as a result, some other concerns may be overlooked. Our quantitative data is simulated by the related secondary data available online, such as Ontario census data, government statistics, and independent research. It is confirmed that the data used is consistent, precise, and reputable. That being said, our research is limited to the data published, leaving our data vulnerable to slight inaccuracy without the knowledge of the survey methodology behind the secondary data. More specifics of the weakness will be demonstrated below. Future studies can be conducted by distributing the survey to desired respondents with more detailed questions aiming to help residents of Ontario improve their situation under COVID impact. 




# Introduction
COVID-19 pandemic has impacted Ontario in various aspects other than just the numbers and cases. Almost everyone has experienced a drastic change in their lives due to the pandemic, and some of the changes, positive or negative, have had an important impact on their political party of choice. In this report, we designed and analyzed the survey “Ontario COVID impact”, intending to discover the influence of COVID-19 on Ontario resident’s political stance and voting intentions. 
Upon deciding our survey methodology, we have taken cost, efficiency, responses, and bias into consideration. More of our methodology will be discussed below. Consequently, the survey consists of socio-demographic, COVID impacts, political stance, and voting intentions, in which we then distributed through phone calls by simple random sampling without replacement (details of the survey attached to appendix). Such probability sampling technique provides a projectable result, and it allows us to apply statistical procedures to analyze data and calculate errors. 
Our findings have revealed that only 53% of respondents consider the public health response “appropriate”, whereas 29% think it is insufficient. We have also discovered COVID pandemic’s impact on different aspects of Ontario residents. In general, the majority of the population is not heavily impacted. However, a significant proportion are undergoing negative impacts from the pandemic in certain aspects such as mental health and financial situation. Despite that 77% of the population are satisfied with the Ontario government’s response, there is still a lot to be done in order to meet the residents’ needs. It is important to take action on such issues as we have discovered 21% of decided voter’s voting intention is influenced by COVID pandemic. Thus the electoral fortune of the Progressive Conservative Party in 2022 is broadly influenced by people’s satisfaction of their response. Our findings have some potential weaknesses and limitations that will later be discussed in detail, as well as some achievable future studies. 

Link to code: https://github.com/1234567980/STA302-PS2


# Methodology/Sampling method
The population being considered consists of all Ontarians aged 18 (years) and above who are legally able to vote. As a result, the Permanent Register of Electors for Ontario (Ontario Voters list) was used as our sampling frame. We then proceeded to sample 3850 Ontarians from the frame, using the SRSWOR method. Although the SRSWOR technique is meant to be used on a sampling frame containing all population units, the Permanent Register of Electors for Ontario contains 94% coverage of our target population. While this does create an opportunity for coverage error to arise, the significance is outweighed by the convenience of using the Ontario Voters list. By randomly selecting members of the Ontario Voters list one at a time (without replacement) to be surveyed over the phone, each Ontarian on the list has an equal probability of being selected. With this survey sampling design, it is unlikely that we would not be able to reach our desired respondents. Since conducting a single survey takes about 3 minutes, and we wish to survey 3850 people, roughly 11,550 minutes, or 192.5 hours, will be required to complete our survey sampling design. Thus, hiring a few employees to direct the surveys at a pay rate of $20 an hour would cost approximately $3850. Assuming that $3850 is the monthly cost of providing polling updates to the Progressive Conservative Party of Ontario, the year would cost $46,200. To deal with non-response, a voicemail will be left to any Ontarian who does not answer the phone the first time they are called.


Below is a preview of our simulated data: 


```{r, echo=FALSE, include=FALSE}
#Load Data

# load cesR package and labelled package
library(labelled)
library(dplyr)
library(ggplot2)
library(readxl)

# Read in Ontario Census Data
OntarioCensus <- read_excel("./OntarioCensus.xlsx")
names(OntarioCensus) <- c("Columns", "Age", "Male", "Female", "NA")

# Rename columns and clean up data to only include data rows that we want
keeps <- c("Columns", "Age", "Male", "Female")
OntarioCensus <- OntarioCensus[keeps]
OntarioCensus <- OntarioCensus[-c(1:26), ]
OntarioCensus <- OntarioCensus[-c(3, 102:105), ]
OntarioCensus <- OntarioCensus[ c(1:7, 8, 14, 20,26,32,38,44,50,56),]
OntarioCensus <- rbind(c("18 to 24 years old",sum(sapply(OntarioCensus[1:7,2], as.numeric)), sum(sapply(OntarioCensus[1:7,3], as.numeric)), sum(sapply(OntarioCensus[1:7,4], as.numeric))), OntarioCensus)
OntarioCensus <- OntarioCensus[-c(2:8), ]

# Calculate the gender percents
total <- sum(sapply(OntarioCensus[1:10,2], as.numeric))
percent_male <- sum(sapply(OntarioCensus[1:10,3], as.numeric)) / total
percent_female <- sum(sapply(OntarioCensus[1:10,4], as.numeric)) / total

# Combine age ranges to suit the answers available in our data
OntarioCensus <- rbind(c("25 to 34 years old",sum(sapply(OntarioCensus[2:3,2], as.numeric)), sum(sapply(OntarioCensus[2:3,3], as.numeric)), sum(sapply(OntarioCensus[2:3,4], as.numeric))), OntarioCensus)
OntarioCensus <- OntarioCensus[-c(3:4), ]
OntarioCensus <- rbind(c("35 to 44 years old",sum(sapply(OntarioCensus[3:4,2], as.numeric)), sum(sapply(OntarioCensus[3:4,3], as.numeric)), sum(sapply(OntarioCensus[3:4,4], as.numeric))), OntarioCensus)
OntarioCensus <- OntarioCensus[-c(4:5), ]
OntarioCensus <- rbind(c("45 to 54 years old",sum(sapply(OntarioCensus[4:5,2], as.numeric)), sum(sapply(OntarioCensus[4:5,3], as.numeric)), sum(sapply(OntarioCensus[4:5,4], as.numeric))), OntarioCensus)
OntarioCensus <- OntarioCensus[-c(5:6), ]
OntarioCensus <- rbind(c("55 years and older",sum(sapply(OntarioCensus[5:7,2], as.numeric)), sum(sapply(OntarioCensus[5:7,3], as.numeric)), sum(sapply(OntarioCensus[5:7,4], as.numeric))), OntarioCensus)
OntarioCensus <- OntarioCensus[-c(6:8), ]

# Add new column for the percentage of each age range
OntarioCensus <- OntarioCensus %>% mutate(Percent = as.numeric(Age) / total)

# Probability weights for survery Q5
# PC, NDP, LIB, GRN
who2018_percentage <- c(0.4063, 0.3369, 0.193, 0.0462, 0.0176)

#TODO: Discuss later? Not sure how to find the appropriate amount of survey takers
survey_size <- 3850

set.seed(123)

# Simulate random sampling from population with our survey (WITHOUT replacement)

my_data <- tibble(

# Probability parameters were calculated above
gender = sample(c('Male', 'Female'), survey_size, replace = TRUE, prob = c(percent_male, percent_female)),
age_range = sample(c('18-24','25-34', '35-44' ,'45-54', '55+'),survey_size, replace = TRUE, prob = sapply(OntarioCensus[1:5, 5], as.numeric)),

who2018 = sample(c('PC', 'NDP', 'LIB', 'GRN','Other'),survey_size, replace = TRUE, prob = who2018_percentage),

# Survey Q6 - Sufficient public health response?
# Source: Angus Reid - Provincial Leadership 
# Too far: 18% | Not far enough: 30% | Just right: 50% | N/A: 2%
q6 = sample(c('Too far', 'Not far enough', 'Just right'), survey_size, replace = TRUE, prob = c(0.18, 0.3, 0.5)),

# Survey Q7 - How has covid19 affected your work?
# Source: Angus Reid - working from home ->QA8 
# Great: 24% | Okay: 61% | Awful: 12% | N/A: 3%
q7 = sample(c('Great', 'Okay', 'Awful'), survey_size, replace =TRUE, prob = c(0.24, 0.61, 0.12)),

# Survey Q8 - Household financial needs
# Source: Angus Reid - Financial Circumstances -> Q9 
# Great: 12% | Good: 63% | Bad: 19% | N/A: 6% 
q8 = sample(c('Great', 'Good', 'Bad'), survey_size, replace = TRUE, prob = c(0.12, 0.63, 0.19)),

# Survey Q9 - Mental Health
# Source: Angus Reid - Covid Concerns + Mental Health -> QA4
# Great: 9% | Good: 53% | Not Good: 33% | N/A: 0.05%
q9 = sample(c('Great', 'Good', 'Bad'), survey_size,replace = TRUE, prob = c(0.09, 0.53, 0.33)),

# Survey Q10 - Satisfaction of provincial gov't response?
# Source: Angus Reid - Provincial Leadership 
# Good Job: 74% | Bad Job: 23% | N/A: 3%
q10 = sample(c('Good', 'Bad'), survey_size, replace = TRUE, prob = c(0.74, 0.23)),

# survey Q11 
# Source: Ontario 2018 election results
q11 =  sample(c("PC", "NDP", 'LIB', 'GRN', 'Other'), survey_size, replace = TRUE, prob = c(0.45, 0.28, 0.22, 0.04, 0.01)),

# Survey Q12
# Yes: 20.9% | No: 74.1% | Others: 5%
q12 = sample(c('Yes', 'No'), survey_size, replace = TRUE, prob = c(0.209, 0.741)),
)

# Preview of data
head(my_data)
```

```{r, echo=FALSE}

# Get a copy the simulate survey results, except re-categorize who they voted for into just PC and not PC
not_PC <- my_data

for (i in 1:dim(not_PC)[1]){
  if (not_PC$who2018[i]!="PC" ){
    not_PC$who2018[i]<- "Not PC"
  }
  
  if (not_PC$q11[i] != "PC"){
    not_PC$q11[i]<- "Not PC"
  }
}

# Calculate the percentage of people who changed from voting for PC to not voting to PC and being negatively affected by Covid19 
# OR from not voting for PC to voting for PC due having a positive experience since during the Covid19 pandemc
num_changed_corona_experience <- 0
for (i in 1:dim(not_PC)[1]){
  if (not_PC$who2018[i]!=not_PC$q11[i]){
    
    if ((not_PC$who2018[i] == 'PC' && (not_PC$q7[i] == 'Awful' || not_PC$q8[i] == 'Bad' || not_PC$q9[i] == 'Bad')) ||
        (not_PC$who2018[i] == 'Not PC' && (not_PC$q7[i] == 'Great' || not_PC$q8[i] == 'Great' || not_PC$q9[i] == 'Great'))
        ){
        num_changed_corona_experience  <- num_changed_corona_experience  + 1;
    }
    
  }
}
# NOTE: This percent was used to simulate survey responses as we added it manually as numbers for the prob parameters
# Since we used my_data to come up with this statistic
percent_changed_corona_experience <- num_changed_corona_experience / nrow(not_PC)

# Get count and percentage of people who voted for PC
voted_PC <- not_PC %>% group_by(who2018) %>% 
    count() %>% mutate(Percent = n / survey_size)

# Get count and percentage of people who did not voted for PC
will_vote_PC <- not_PC %>% group_by(q11) %>% 
    count() %>% mutate(Percent = n / survey_size)

# Combine previous 2 data frames into one for graphing
comparisons_PC <- data.frame("Labels" = c("Voted for PC (2018)", "Will vote for PC"), "Percentage" = c(as.numeric(voted_PC[2,3]), as.numeric(will_vote_PC[2,3])))


```



# Discussion and Results

With the social distancing regulations and mask requirements that were put into place, there has been much debate online about whether these measures were going too far or not enough. From Figure (1), we've noticed most Ontarians think that the Public Health's approach to Corona was appropriate. However, it is interesting to note that 29% of people think that the government's response to the current pandemic was not sufficient in comparison to 18% of people who think that regulations have gone too far. In addition to the results found in this survey, a nation-wide survey confirmed that the government satisfaction rate in April in Ontario was 78%. However, it is important to note that this number has significantly dropped since then. The same survey(bib2) was conducted in September and the satisfaction rate in Ontario sank to 63%. The decrease in satisfaction rate in September could probably be attributed to the rise of Covid19 cases that we are currently seeing.



```{r, echo=FALSE}
# Bar plot of the answer distribution of q6
#Scale fill manual not matching the name as we don't want the color to show on the side, but still manually fill the color
ggplot(my_data,aes(x = factor(q6, level = c("Not far enough", "Just right", "Too far", "N/A")), fill=q6, label = scales::percent(prop.table(stat(count))))) +
    geom_bar()+
    scale_x_discrete("" ,labels = c("Just right" = "Appropriate", "Not far enough" = "Insufficient" ,
                                "Too far"= "Too Extreme", "N/A"="Prefer Not To Say"))+
  scale_y_continuous(limits = c(0,2200))+
  scale_fill_manual("legend",breaks = c("A", "B", "C", "D"), values =
                        c("Just right" = "dodgerblue1", "Not far enough" = "dodgerblue4", "Too far"="skyblue1", "N/A"="grey"))+ 
  geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 4)+
  labs(title = "Opinion on the public health response to \n the current coronavirus pandemic in Ontario", tag = "(1)")+
  theme(plot.title = element_text(lineheight=1, face="bold"))+
  theme_light()+
  theme(axis.title.y=element_blank(),
        axis.text.y =element_blank(),
        axis.ticks.y =element_blank())+
  theme(axis.text.x= element_text(colour = "black", size = 12))
```


From small business owners to those working in the gig economy and artists, the government response has been lackluster as a CBC article states that business support from the government “is not even close to being enough” (bib1). 



```{r, echo=FALSE}
# Bar plot of the answer distribution of q7
ggplot(my_data,aes(x = factor(q7, level = c("Awful", "Okay", "Great")), fill=q7, label = scales::percent(prop.table(stat(count))))) +
    geom_bar()+
    scale_x_discrete("" ,labels = c("Awful" = "Impaired My Work", "Okay" = "No Change in My Work" ,
                                "Great"= "Improved My Work"))+
    scale_y_continuous(limits = c(0,3000))+
    scale_fill_manual(breaks = c("Impaired My Work", "No Change in My Work", "Improved My Work"),values =
                        c("Awful" = "dodgerblue4", "Okay" = "dodgerblue1", "Great"="skyblue1"))+ 
  geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 4)+
  labs(title = "How has the coronavirus pandemic affected your work", tag = "(2)")+
  theme(plot.title = element_text(lineheight=1, face="bold"))+
  theme_light()+
  theme(axis.title.y=element_blank(),
        axis.text.y =element_blank(),
        axis.ticks.y =element_blank())+
  theme(axis.text.x= element_text(colour = "black", size = 12))
```



Moreover, since many people have been forced to work from home, many complaints have emerged about the difficulty of working in a home environment. However, it is possible that unhappy workers were more vocal about such issues, as 64% of the respondents indicate that they have not been heavily impacted by COVID-19 in graph (2). And 24% have even seen positive effects from the pandemic.



Since the beginning of the pandemic, finances have been a major point of concern due to the disruption caused by the virus in many industries. It was reported that Canada’s real GDP as an annualized rate fell by  38.7%. Statistics Canada reported that household expenses dropped by 13.1% signifying that the pandemic has had a significant impact on families in Canada. However, both federal and provincial governments have offered the aid of some sort to help offset the financial burden on affected families. 

```{r, echo=FALSE}
# Bar plot of the answer distribution of q8
ggplot(my_data,aes(x = factor(q8, level = c("Bad", "Good", "Great")), fill=q8, label = scales::percent(prop.table(stat(count))))) +
    geom_bar()+
    scale_x_discrete("" ,labels = c("Bad" = "More Difficult", "Good" = "Neither More Difficult Nor Easier" ,
                                "Great"= "Easier"))+
    scale_y_continuous(limits = c(0,2700))+
    scale_fill_manual(breaks = c("More Difficult", "Neither More Difficult Nor Easier", "Easier"),values =
                        c("Bad" = "dodgerblue4", "Good" = "dodgerblue1", "Great"="skyblue1"))+ 
  geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 4)+
  labs(title = "Since the coronavirus pandemic, \n how has it been for you and your household to meet your financial needs", tag = "(3)")+
  theme(plot.title = element_text(lineheight=1, face="bold"))+
  theme_light()+
  theme(axis.title.y=element_blank(),
        axis.text.y =element_blank(),
        axis.ticks.y =element_blank())+
  theme(axis.text.x= element_text(colour = "black", size = 12))
```


From graph (3), it seems that for a majority of people, the coronavirus pandemic has not had a major impact on their finances with even a small amount of people reporting that their finances have benefited from the pandemic. Suggesting that the emergency assistance provided by the government has been successful to some degree. However, it is still concerning to notice that a sizable portion of the population, 22.4%, has reported that their financial situations have become more difficult since the start of the pandemic. 




Additionally, a study by iPolitics found that financial difficulties played a role in the increase of mental health struggles as distress calls went up in August due to anxieties of the termination of CERB (bib3). This finding complements our survey’s results as 35% of people reported feeling that their mental health has gotten worse from before the pandemic. This result is confirmed by an additional survey conducted by CTV measuring the rates of anxiety/depression diagnoses during the pandemic. “Nearly 50 percent were considered probable candidates for anxiety disorders and more than 40 percent were likely to be clinically depressed. Almost 85 percent of respondents reported moderate to high stress(bib4).” In summary, “there has been a lot of pressure on people’s relationships, a lot of pressure on people’s living arrangements, and of course - the financial pressures (bib5).”


```{r, echo=FALSE,  include=FALSE}
# Bar plot of the answer distribution of q9
ggplot(my_data,aes(x = factor(q9, level = c("Bad", "Good", "Great")), fill=q9, label = scales::percent(prop.table(stat(count))))) +
    geom_bar()+
    scale_x_discrete("" ,labels = c("Bad" = "Worse", "Good" = "About the same" ,
                                "Great"= "Better"))+
    scale_y_continuous(limits = c(0,3500))+
    scale_y_continuous(limits = c(0,2500))+
    scale_fill_manual(breaks = c("Worse", "About the same", "Better"),values =
                        c("Bad" = "dodgerblue4", "Good" = "dodgerblue1", "Great"="skyblue1"))+ 
  geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 4)+
  labs(title = "Compared to before the coronavirus pandemic,\n how would you rate your mental health now", tag = "(4)")+
  theme(plot.title = element_text( face="bold"))+
  theme_light()+
  theme(axis.title.y=element_blank(),
        axis.text.y =element_blank(),
        axis.ticks.y =element_blank())+
  theme(axis.text.x= element_text(colour = "black", size = 12))
```



So far it seems that for a majority of people, the coronavirus pandemic has not impacted them heavily in a detrimental way. This is reflected by 77% of our respondents responding that they were satisfied with how the provincial government has handled the pandemic, shown in graph (5). It demonstrates that although the provincial government has been successful at preventing many people from being negatively influenced by COVID-19, there is still much room for improvement with 23% of the population unsatisfied. 



```{r, echo=FALSE}
# Bar plot of the answer distribution of q10
ggplot(my_data,aes(x = factor(q10, level = c("Bad", "Good")), fill=q10, label = scales::percent(prop.table(stat(count))))) +
    geom_bar(width = 0.6)+
    scale_x_discrete("" ,labels = c("Bad" = "Dissatisfied", "Good" = "Satisfied"))+
    scale_y_continuous(limits = c(0,3500))+
    scale_fill_manual(breaks = c("Dissatisfied", "Satisfied"),values =
                        c("Bad" = "dodgerblue4", "Good"="skyblue1"))+ 
  geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 4)+
  labs(title = "How satisfied are you with the provincial government's \n response to the COVID-19 pandemic", tag = "(5)")+
  theme(plot.title = element_text( lineheight=1, face="bold"))+
  theme_light()+
  theme(axis.title.y=element_blank(),
        axis.text.y =element_blank(),
        axis.ticks.y =element_blank())+
  theme(axis.text.x= element_text(colour = "black", size = 12))
```



Graph (6) depicts the percentage of the population that would vote for a particular party if a sudden election were to be held right now. The graph above demonstrates that the largest portion of the population that would vote for the Progressive Conservitive Party corresponding to 79% of the population is satisfied with the provincial government’s response(figure (5)). Figures 7 further enforces that the Ontario government has responded in the right direction by showing that the percentage of people who would vote for Conservatives has increased by 5.5% from 2018. 


```{r, echo=FALSE}
# Bar plot of the answer distribution of q11
ggplot(my_data,aes(x = factor(q11, level = c("PC", "NDP", "LIB", "GRN", "Other")), fill=q11, label = scales::percent(prop.table(stat(count))))) +
    geom_bar()+
    scale_x_discrete("" ,labels = c("PC" = "Progressive \n Conservative \n Party", "NDP" = "New Democratic \n Party", "LIB"= "Liberal Party", "GRN"="Green Party", "Other"="Other"))+
    scale_y_continuous(limits = c(0,2000))+
    scale_fill_manual(breaks = c("A", "B", "C", "D","E"), values = c("PC" = "dodgerblue1", "NDP" = "orange", "LIB"= "red", "GRN"="limegreen", "Other"="grey"))+ 
  geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 4)+
  labs(title = "Vote Intention in Ontario(Decided Voters)", tag = "(6)")+
  theme(plot.title = element_text(lineheight=1, face="bold"))+
  theme_light()+
  theme(axis.title.y=element_blank(),
        axis.text.y =element_blank(),
        axis.ticks =element_blank())+
  theme(axis.text.x = element_text(color="black", 
                           size=12))
```



```{r, echo=FALSE}
# Bar plot of a comparison between the people who voted PC in 2018 and the people who would vote for PC now
ggplot(comparisons_PC, aes(x = Labels, y = Percentage, fill = Labels  , label = paste(round(Percentage*100, 1),"%",sep=""))) + 
  geom_bar(width = 0.6, stat = "identity") +
  geom_text(position = position_dodge(.9), 
              vjust = -0.5, 
              size = 4) +
  scale_x_discrete("", labels = c("Voted for Conservatives \n in 2018 Election", "Would vote for Conservatives \n today(Decided voters)")) +
  coord_cartesian(ylim = c(0,0.6))+
  scale_fill_manual(breaks = c("Voted for Conservatives in 2018 Election", "Would vote for Conservatives today(Decided voters"), values = c("Voted for PC (2018)" = "dodgerblue4", "Will vote for PC"="skyblue1"))+
  labs(title = "Shares of Votes in 2018 Provincial Election vs. \n Decided Voters Today", tag = "(7)")+
  theme(plot.title = element_text(hjust = 0.5 , lineheight=1, face="bold", colour = "black"))+
  theme_light()+
    theme(axis.text.x= element_text(colour = "black", size = 12))+
  theme(axis.title.y=element_blank(),
        axis.text.y =element_blank(),
        axis.ticks.y =element_blank())
```




```{r, echo=FALSE, include=FALSE}
# Bar plot of the answer distribution of q12
ggplot(my_data,aes(x = factor(q12, level = c('Yes', 'No')), fill=q12, label = scales::percent(prop.table(stat(count))))) +
    geom_bar(width = 0.6)+
    scale_x_discrete("" ,labels = c('Yes' = "Influenced by \n government response", 'No' = "Not influenced \n by government response"))+
    coord_cartesian(ylim=c(0,3500))+
    scale_fill_manual(breaks = c("Yes", "No"),values =
                        c('Yes' = "dodgerblue1", 'No' = "dodgerblue4"))+ 
  geom_text(stat = 'count',
              position = position_dodge(.9), 
              vjust = -0.5, 
              size = 4)+
  labs(title = "Is your vote influenced by the government's response \n to the COVID 19 pandemic", tag = "(8)")+
  theme(plot.title = element_text(lineheight=1, face="bold"))+
  theme_light()+
  theme(axis.title.y=element_blank(),
        axis.text.y =element_blank(),
        axis.ticks.y =element_blank())+
  theme(axis.text.x= element_text(colour = "black", size = 12))
```



As shown from the graph above (Figure(8)), it seems that for most people, their political stance has not been severely influenced by the Ontario government’s response. Meaning that only approximately 21% of people changed their stances from supporting or not supporting the PC party due to being negatively or positively affected by the Covid19 pandemic, respectively. The small increase in the percentage of the population who support the PCC reveals that just a satisfactory response is not enough to sway voters.

Overall, it seems that the current Ontario government’s response has been alright and enough to keep their current voters content. However, with 35% of the population reporting a decline in their mental health, it is clear that there are many underlying issues that can and should be further addressed. There is much room for improvement that the PCC should pay attention to if the they aims to capture more votes.



# Limitations

The main topic of concern in our survey is the COVID-19 pandemic which is currently ongoing, as a result there were many limitations to the data that we could find for estimating the survey responses. 

The choices that were available on the survey did not always match the data that we could find. Thus, some options were grouped together into the same section during the survey simulation. Moreover, N/A choices were not always included in the probabilities that could be found and as a result we had to manually scale the probabilities.

he data obtained to determine the probability of each answer were obtained from various sources. Moreover, many of these sources had different survey methodologies that were not entirely divulged to the public. Meaning that we are unable to determine any inaccuracies due to lack of information.

Moreover, many of these sources did not disclose their entire dataset. As a result, each variable was simulated independently based on just the reported probabilities. Thus, although we can get a clear picture of the variable independently, we are unable to determine how the variables influence each other.

In addition, we can report an estimate of how the PC government is doing overall because our survey aims to survey the general feelings of the overall Ontario population. However, since the survey targets such broad concerns, it is difficult to figure out specifics of where to improve. From our analysis, we’ve found that 22% of people have had their financial situations negatively impacted by the pandemic. However, it is difficult to tell whether the larger concern is the employment rate, retirement savings, or something else. 

It is also difficult to determine answers to questions such as which industry is being affected the most causing people to lose their jobs using only the results from the survey. Furthermore, a majority of the survey answers options rely on self-perception, which can vary by person. Thus the results might be inexact. 


# Next Steps
The second wave has already caused an increase in public uncertainty and is disrupting the approval rating of the PC party. Government policymakers must monitor the situation for the need to tighten restrictions and put increased effort in improving their approach to containing the impact of COVID-19. 



# Citations

1. Allaire, JJ, et al. Rmarkdown: Dynamic Documents for R. 2020, github.com/rstudio/rmarkdown. Accessed 9 Oct. 2020. R package version 2.4.

2. “As Pandemic Endures, Three-in-Ten Canadians Say Restrictions in Their Own Province Don’t Go Far Enough.” Angus Reid Institute, 13 Aug. 2020, angusreid.org/covid-19-provincial-leadership/. Accessed 10 Oct. 2020.

3. “Canadians’ Personal Financial Circumstances Improving, but Majority Will Defer Major Purchases in the next Year.” Angus Reid Institute, 23 July 2020, angusreid.org/covid-19-finances-spending/. Accessed 10 Oct. 2020.

4. Canseco, Mario. “Satisfaction with COVID-19 Handling Drops in Ontario and Quebec.” Research Co., 19 May 2020, researchco.ca/2020/05/19/canada-governments-covid19/. Accessed 10 Oct. 2020.

5. Government of Canada, Statistics Canada. “2016 Census Data Tables – Age (in Single Years) and Average Age (127) and Sex (3) for the Population of Canada, Provinces and Territories, Census Divisions, Census
Subdivisions and Dissemination Areas, 2016 Census - 100% Data.” Statcan.Gc.Ca, 2016, www12.statcan.gc.ca/census-recensement/2016/dp-pd/dt-td/Rp-eng.cfm?TABID=2&LANG=E&APATH=3&DETAIL=0&DIM=0&FL=A&FREE=0&GC=0&GID=0&GK=0&GRP=1&PID=109525&PRID=35&PTYPE=109445&S=0&SHOWALL=0&SUB=0&Temporal=2016&THEME=115&VID=0&VNAMEE=&VNAMEF=&D1=0&D2=0&D3=0&D4=0&D5=0&D6=0. Accessed 10 Oct. 2020.

6. Hadley Wickham. Ggplot2 Elegant Graphics for Data Analysis. Cham Springer International Publishing, 2016.
Larmarange, Joseph. Labelled: Manipulating Labelled Data. 2020, larmarange.github.io/labelled/. package version 2.7.0.

7. Oct 8, Janet E. Silver, and 2020 6:21pm. “As COVID Cases Rise, so Do Calls for More Mental-Health Funding.” IPolitics, 8 Oct. 2020, ipolitics.ca/2020/10/08/as-covid-cases-rise-so-do-calls-for-more-mental-health-funding/. Accessed 10 Oct. 2020.

8. “Ontario Election Results From CBC News.” CBC News, newsinteractives.cbc.ca/onvotes/results/.

9. “Ontario Newsroom | Salle de Presse de l’Ontario.” News.Ontario.Ca, 30 Sept. 2020, news.ontario.ca/en/release/58602/ontario-releases-updated-covid-19-modelling-for-second-wave. Accessed 10 Oct. 2020.

10. “Ontario Spotlight: As COVID-19 Cases Climb, Can the Ford Government Maintain the Political Goodwill It’s Earned?” Angus Reid Institute, 18 Sept. 2020, angusreid.org/ontario-government-august-2020/. Accessed 10 Oct. 2020.

11. Ore, Jonathan. “Mixed Reception to Feds’ COVID-19 Response among Canadians as Pandemic Marches on | CBC Radio.” CBC, 12 July 2020, www.cbc.ca/radio/checkup/what-grade-do-you-give-the-trudeau-government-on-covid-19-1.5643270/mixed-reception-to-feds-covid-19-response-among-canadians-as-pandemic-marches-on-1.5646508. Accessed 10 Oct. 2020.

12. Press, Jordan. “Canadian Economy Posted Steepest Decline on Record as Coronavirus Struck: StatCan.” CTVNews, 28 Aug. 2020, www.ctvnews.ca/business/canadian-economy-posted-steepest-decline-on-record-as-coronavirus-struck-statcan-1.5082814. Accessed 10 Oct. 2020.

13. R Core Team. R: A Language and Environment for Statistical Computing. Vienna, Austria, R Foundation for Statistical Computing, 2020, www.R-project.org/. Accessed 9 Oct. 2020
“Second Wave Angst: COVID-19 Concern Levels Rebound to April Highs as Canadians Brace for Worse to Come.” Angus Reid Institute, 30 Sept. 2020, angusreid.org/covid-second-wave-september/. Accessed 10 Oct. 2020.

14. Szperling, Peter. “Anxiety Levels Increasing during COVID-19 Pandemic: Study.” CityTV Ottawa, 24 Sept. 2020, ottawa.ctvnews.ca/anxiety-levels-increasing-during-covid-19-pandemic-study-1.5119162. Accessed 8 Oct. 2020.

15. “So Long, Office Space? Two-Thirds of Canadians Who Work from Home Expect It to Continue after Pandemic.” Angus Reid Institute, 11 June 2020, angusreid.org/coronavirus-work-from-home/. Accessed 10 Oct. 2020.

16. Weber, Bob. “COVID-19 Causing Stress, Depression and Obsessive Behaviour: Survey.” CTV News, 26 Sept. 2020. Accessed 8 Oct. 2020.

17. Wickham, Hadley, et al. Dplyr: A Grammar of Data Manipulation. 2020, dplyr.tidyverse.org, github.com/tidyverse/dplyr. Accessed 9 Oct. 2020.

18. Wickham, Hadley, and Jennifer Bryan. Readxl: Read Excel Files. 2019, readxl.tidyverse.org, github.com/tidyverse/readxl.

19. Yihui Xie, et al. R Markdown Cookbook.. Chapman and Hall/CRC, 2020.

20. Yihui Xie, et al. R Markdown: The Definitive Guide.. Chapman and Hall/CRC, 2018.

21. Canada, Elections. “Description of the National Register of Electors.” Www.Elections.Ca, 7 July 2020, www.elections.ca/content.aspx?section=vot&dir=reg/des&document=index&lang=e. Accessed 9 Oct. 2020.


# Appendix
A screen shot is not avaliable due to survey monkey limitations. 
Link to survey at:
https://www.surveymonkey.com/create/preview/?sm=grlkLOM6Xugn6hymzrWHCoMlQNw7TDU4F1YeaxoK76I_3D&tab_clicked=1
